Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-19 paper.tex
eval everywhere

\newcommand{\judgeCeval}[2]{#1 \Downarrow_{\mathcal{C}} #2}

@@

\newcommand{\judgeSeval}[3]{#2 \Downarrow_{\mathcal{S}(#1)} #3}

@@

\newcommand{\judgeDeval}[2]{#1 \Downarrow_{\mathcal{D}} #2}

@@

@@

Many of the existing research on coercions or threesomes include proof(s) of 
correctness, that is, the cast calculus reduces to an observation ($ e 
\Downarrow o $) if and only if their space-efficient calcului reduce to the 
same observation ($ e \Downarrow_{*} o $), where an observation is either a 
value constructor or 
blame. 
Despite the striking similarity of these proofs, they are all built from ground 
up.

@@

\redrule{
\sOOhalt{o}}{
\sOOhalt{o}}{}
	\end{array}
	\]	
%	
%	\fbox{$case(v, \kappa, \kappa) = s$}
%	\[
%	\begin{array}{rclr}
%	\funrule{case(\vOOinl{v},\kappa_1,\kappa_2)}{
%		\sOOreturn{v}{\kappa_1}
%	}{}
%	\funrule{case(\vOOinr{v},\kappa_1,\kappa_2)}{
%		\sOOreturn{v}{\kappa_2}
%	}{}
%	
%\funrule{case(\vOOcast{v}{\cOOcast{\POOsum{T_{11}}{T_{12}}}{l}{\POOsum{T_{21}}{T_{22}}}},\kappa_1,\kappa_2)}{
%		case(v,
%		\langle \cOOcast{T_{11}}{l}{T_{21}} \rangle \kappa_1,
%		\langle \cOOcast{T_{12}}{l}{T_{22}} \rangle \kappa_2)
%	}{}
%	\end{array}
%	\]

@@

The reflexive transitive closure of reduction ($ \judgeCreduceTrans{s}{s} $) 
and evaluation ($ \judgeCeval{e}{o} $) are standard.

@@

In subsection~\ref{ssec:framework:cek}, we introduce \effCEK{C}, a family 
of space-efficient semantics indexed by Cast ADT ($ C $). When $ C $ is a Lazy 
D Cast ADT, \effCEK{C} repects \ineffCEKD.
In subsection~\ref{secc:framework:monoid-correct}, we prove when $ C $ is
a Lazy D Cast ADT and a monoid as well, $ \judgeDeval{e}{o} $ if 
and only if $ \judgeSeval{C}{e}{o} $. In 
subsection~\ref{secc:framework:all-correct}, we further prove that when $ C $ is
a Lazy D Cast ADT, $ \judgeDeval{e}{o} $ if and only if $ \judgeSeval{C}{e}{o} 
$.

@@

In this subsection, we prove that if a instance of Cast ADT $ C $ is Lazy D and 
is a monoid, then $ \judgeDeval{e}{o} $ if and only if $ \judgeSeval{C}{e}{o} 
$. We proof this theorem with a weak bisimulation between \ineffCEKD\ and 
\effCEK{C}.

@@

\begin{lemma}[$ \text{Lazy D Cast ADT} \cap \text{Monoid} $ Respects 
	$ \mathcal{C} $]
	\label{thm:surely-monoidic-eval}
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ and $ C $ is a 
	\lazyD{} Cast ADT and a monoid,
	\[
	\judgeSeval{C}{e}{o}
	 \; \text{if and only if} \; 
	\judgeDeval{e}{o}
	\]
\end{lemma}
\begin{proof}Proving left implies right is trivial.
To prove $ \judgeDeval{e}{o} $ implies $ \judgeSeval{C}{e}{o} $, 
we firstly take one steps in \effCEK{C}, and catch it up in \ineffCEKD\ by 
taking one or more steps in \ineffCEKD. As we just took a positive number of 
steps, we are closer to the halting state.
\end{proof}

@@

In this subsection, we prove that for all $ C $, if $ C $ is a Lazy D Cast ADT, 
then $ \judgeDeval{e}{o} $ if and only if $\judgeSeval{C}{e}{o}$. We first 
prove that all $ S(C) $ where $ C $ is a Lazy D Cast ADT are equivalent, then 
connect this theorem to Lemma~\ref{thm:surely-monoidic-eval}.

@@

\begin{proposition}[Equivalence of \lazyD{} Cast ADT]
	\label{thm:surely-lazyD-eqv}
	If $ \judgetype{\emptyset}{e}{T} $, $ o : T $, and $ C_1 $ and $ C_2 $ 
	are \lazyD\ Cast ADTs,
	\[
	e \Downarrow_{S(C_1)} o \; \text{if and only if} \; 
	e \Downarrow_{S(C_2)} o
	\]
\end{proposition}
\begin{proof}
	By induction with the help of Lemma~\ref{thm:CEKS-bisim}.
\end{proof}

\begin{theorem}[\lazyD\ Cast ADT Respects \ineffCEKD]
	\label{thm:surely-lazyD-correct}
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ and $ C $ is a 
	\lazyD{} Cast ADT.
	\[
	\judgeCeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{C}{e}{o}
	\]
\end{theorem}
\begin{proof}
	If we have a surely LazyD Cast ADT and it is a monoid, this 
	proof is immediately from Lemma~\ref{thm:surely-monoidic-eval} and 
	Proposition~\ref{thm:surely-lazyD-eqv}. One such representation is to 
	represent casts as a list of triples of type, label, and type, where $ 
	id(T) $ is the empty list and $ seq(c,c,) $ is list append. Another such 
	representation is the \lazyD\ hypercoercion!
\end{proof}

@@

\begin{theorem}[\lazyD{} hypercoercion Respect \ineffCEKD]
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ 
	\[
	\judgeCeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{H}{e}{o}
	\]
\end{theorem}
\begin{proof}
	Immediately from Theorem~\ref{thm:surely-lazyD-correct} and 
	Lemma~\ref{thm:hc-surely-lazyD}.
	Alternatively, from Lemma~\ref{thm:surely-monoidic-eval},
	Lemma~\ref{thm:hc-surely-lazyD}, and 
	Proposition~\ref{thm:hc-monoid}.
\end{proof}

>>>>

\newcommand{\judgeCeval}[2]{eval_{\mathcal{C}}(#1) = #2}

@@

\newcommand{\judgeSeval}[3]{eval_{\mathcal{S}(#1)}(#2) = #3}

@@

\newcommand{\judgeDeval}[2]{eval_{\mathcal{D}}(#1) #2}

@@

\newcommand{\evalEqv}[2]{$ eval_{\text{#1}} = eval_{\text{#2}} $}

@@

Many of the existing research on coercions or threesomes include proof(s) of 
correctness, that is, the evaluation (partial) function of cast calculus $ eval 
$ is equal to their space-efficient evaluation ($ eval_{*} $).
Despite the striking similarity of these proofs, they are all built from ground 
up.

@@

%\redrule{
%\sOOhalt{o}}{
%\sOOhalt{o}}{}
	\end{array}
	\]	
%	
%	\fbox{$case(v, \kappa, \kappa) = s$}
%	\[
%	\begin{array}{rclr}
%	\funrule{case(\vOOinl{v},\kappa_1,\kappa_2)}{
%		\sOOreturn{v}{\kappa_1}
%	}{}
%	\funrule{case(\vOOinr{v},\kappa_1,\kappa_2)}{
%		\sOOreturn{v}{\kappa_2}
%	}{}
%	
%\funrule{case(\vOOcast{v}{\cOOcast{\POOsum{T_{11}}{T_{12}}}{l}{\POOsum{T_{21}}{T_{22}}}},\kappa_1,\kappa_2)}{
%		case(v,
%		\langle \cOOcast{T_{11}}{l}{T_{21}} \rangle \kappa_1,
%		\langle \cOOcast{T_{12}}{l}{T_{22}} \rangle \kappa_2)
%	}{}
%	\end{array}
%	\]

@@

The reflexive transitive closure of reduction ($ \judgeCreduceTrans{s}{s} $) 
and evaluation ($ eval_\mathcal{C}(e) $) are standard.

@@

In subsection~\ref{ssec:framework:cek}, we introduce \effCEK{C}, a family 
of space-efficient semantics indexed by Cast ADT ($ C $). When $ C $ is a Lazy 
D Cast ADT, \effCEK{C} repects \ineffCEKD.
In subsection~\ref{secc:framework:monoid-correct}, we prove when $ C $ is
a Lazy D Cast ADT and a monoid as well, $ \judgeDeval{e}{o} $ if 
and only if $ \judgeSeval{C}{e}{o} $. In 
subsection~\ref{secc:framework:all-correct}, we further prove that when $ C $ is
a Lazy D Cast ADT, \evalEqv{\ineffCEKD}{\effCEK{C}}.

@@

In this subsection, we prove that if a instance of Cast ADT $ C $ is Lazy D and 
is a monoid, then $ eval_\mathcal{D} = eval_{\mathcal{S}(C)} $. We proof this 
theorem with a weak bisimulation between \ineffCEKD\ and 
\effCEK{C}.

@@

\begin{lemma}[$ \text{Lazy D Cast ADT} \cap \text{Monoid} $ Respects 
	$ \mathcal{C} $]
	\label{thm:surely-monoidic-eval}
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ and $ C $ is a 
	\lazyD{} Cast ADT and a monoid,
	\[
	\text{\evalEqv{\effCEK{C}}{\ineffCEKD}}
	\]
\end{lemma}
\begin{proof}Proving left implies right is trivial.
To prove $ \judgeDeval{e}{o} $ implies $ \judgeSeval{C}{e}{o} $, 
we firstly take one steps in \effCEK{C}, and catch it up in \ineffCEKD\ by 
taking one or more steps in \ineffCEKD. As we just took a positive number of 
steps, we are closer to the halting state.
\end{proof}

@@

In this subsection, we prove that for all $ C $, if $ C $ is a Lazy D Cast ADT, 
then \evalEqv{\ineffCEKD}{\effCEK{C}}. We first 
prove that all $ S(C) $ where $ C $ is a Lazy D Cast ADT are equivalent, then 
connect this theorem to Lemma~\ref{thm:surely-monoidic-eval}.

@@

\begin{proposition}[Equivalence of \lazyD{} Cast ADT]
	\label{thm:surely-lazyD-eqv}
	If $ \judgetype{\emptyset}{e}{T} $, $ o : T $, and $ C_1 $ and $ C_2 $ 
	are \lazyD\ Cast ADTs,
	\[
	eval_{S(C_1)} = eval_{S(C_2)}
	\]
\end{proposition}
\begin{proof}
	By induction with the help of Lemma~\ref{thm:CEKS-bisim}.
\end{proof}

\begin{theorem}[\lazyD\ Cast ADT Respects \ineffCEKD]
	\label{thm:surely-lazyD-correct}
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ and $ C $ is a 
	\lazyD{} Cast ADT.
	\[
	\text{\evalEqv{\ineffCEKD}{\effCEK{C}}}
	\]
\end{theorem}
\begin{proof}
	If we have a Lazy D Cast ADT and it is a monoid, this 
	proof is immediately from Lemma~\ref{thm:surely-monoidic-eval} and 
	Proposition~\ref{thm:surely-lazyD-eqv}. One such representation is to 
	represent casts as a list of triples of type, label, and type, where $ 
	id(T) $ is the empty list and $ seq(c,c) $ is the list append. Another such 
	representation is the \lazyD\ hypercoercion!
\end{proof}

@@

\begin{theorem}[\lazyD{} hypercoercion Respect \ineffCEKD]
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ 
	\[
	\text{\evalEqv{\ineffCEKD}{\effCEK{H}}}
	\]
\end{theorem}
\begin{proof}
	Immediately from Theorem~\ref{thm:surely-lazyD-correct} and 
	Lemma~\ref{thm:hc-surely-lazyD}.
	Alternatively, from Lemma~\ref{thm:surely-monoidic-eval},
	Lemma~\ref{thm:hc-surely-lazyD}, and 
	Proposition~\ref{thm:hc-monoid}.
\end{proof}
